Ages of Planets and Stars Observed by Kepler

### 2024-04-03 paper/ms.tex
update ms

Figure~\ref{fig:hist_tgyro} shows the gyrochone age distributions
of stars and planet-hosts in the Kepler field.
We truncated the plots near $3$\,Gyr due to arguments that suggest the
detectability of rotation signals is near-unity up to this age for Sun-like
stars, and decreases for older ages \citep{2022ApJ...937...94M}.
Both the stellar and planetary samples point toward a star
formation rate in the Kepler field that has slowed over the past
few gigayears.
Imagine we were to group stars between 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr, and 
label them ``young'', ``intermediate-age'', and ``old''.
A simple counting exercise from these distributions tells us that there are \ratioobtoybstars\ times as many 
old stars in the Kepler field as young stars.
Similarly, there are \ratioobtoybplanets\ times as many old planet hosts as
young planet hosts.

{\it LGB pending words: touch on findings on MW SFH from chemistry,
isochrones, and asteroseismology}.

%e.g. hist_samples_field_gyro_ages_hist_field_gyro_ages_20230529.pdf

@@

>>>>

Figure~\ref{fig:hist_tgyro} shows the gyrochone age distributions of
stars and planet-hosts in the Kepler field.
We truncated the plots near $3$\,Gyr due to arguments that suggest the
detectability of rotation signals is near-unity up to this age for
Sun-like stars, and decreases for older ages
\citep{2022ApJ...937...94M}.
Both the stellar and planetary samples point toward a star formation
rate in the Kepler field that has slowed over the past few gigayears.
{\it LGB pending words: allude to findings on MW SFH from chemistry,
isochrones, and asteroseismology}.

@@

Imagine we were to group stars between 0-1\,Gyr, 1-2\,Gyr, and
2-3\,Gyr, and label them ``young'', ``intermediate-age'', and ``old''.
A simple counting exercise from the distributions in
Figure~\ref{fig:hist_tgyro} tells us that there are \ratioobtoybstars\
times as many old stars in the Kepler field as young stars.
Similarly, there are \ratioobtoybplanets\ times as many old planet
hosts as young planet hosts.

In terms of actual detected planet counts, our rotation-based ages for
the Kepler sample yield \{\nplyounggyro, \nplmidgyro, \nploldgyro\}
detected planets in the 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr bins.
For the sub-gigayear bin in particular, which holds specific weight in
studies of planet evolution, while there are \nplyounggyro\ planets
with median ages below 1\,Gyr, if we were to require $t_{\rm
gyro}$$<$1\,Gyr at 2$\sigma$ confidence, this would leave only
\nplyounggyrotwosigma\ planets orbiting \nplhostsyounggyrotwosigma\
stars.

\subsection{Sanity Checks}
To date, four stellar clusters in the Kepler field have fielded planet
detections.
They are:...
Young Kepler systems with the most precisely measured ages are...

(TODO: perhaps easiest in a little table.  Theia316, Theia520, CepHer,
and NGC6811).
